Supply diversity

### 2023-04-20 ms_figures.Rmd
Update results with new text

```{r ldc_net_exports_habitat_method}
ldc_net_habitat_method_summary <- read.csv(file.path(datadir, "ldc_net_habitat_method_summary.csv"))

ldc_net_exports_by_source_plot <- ldc_net_habitat_method_summary %>%
  ggplot(aes(x = year, y = net_export/1000000, color = habitat_method)) +
  geom_line() +
  scale_color_manual(values = habitat_source_colors) +
  labs(x = "", y = "Net export (million t)", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

dom_source_ts_p <- dom_source_ts %>%
  filter(dom_source != "error export") %>%
  ggplot(aes(x = year, y = live_weight_t / 1000000, color = dom_source)) +
  geom_line() +
  scale_color_manual(values = c("#D38F35", "#114F59", "grey")) +
  theme_bw() +
  labs(x = "", y = "Live weight\nmillion tonnes", color = "Source") +
  theme(
    legend.position = "bottom",
    legend.box.margin=margin(t = -25)
  )

>>>>

```{r ldc_net_exports_habitat_method}
ldc_net_habitat_method_summary <- read.csv(file.path(datadir, "ldc_net_habitat_method_summary.csv")) %>%
  select(year, habitat_method, net_export)

ldc_net_total <- ldc_net_habitat_method_summary %>%
  group_by(year) %>%
  summarise(net_export = sum(net_export)) %>%
  mutate(habitat_method = "total")

ldc_net_habitat_method_summary <- ldc_net_habitat_method_summary %>%
  bind_rows(ldc_net_total)

ldc_net_exports_by_source_plot <- ldc_net_habitat_method_summary %>%
  ggplot(aes(x = year, y = net_export/1000000, color = habitat_method)) +
  geom_line(size = 1) +
  scale_color_manual(values = habitat_source_colors) +
  labs(x = "", y = "Net export (million t)", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

dom_source_ts_p <- dom_source_ts %>%
  ggplot(aes(x = year, y = live_weight_t / 1000000, color = dom_source)) +
  geom_line() +
  scale_color_manual(values = c("#D38F35", "grey40", "#114F59")) +
  theme_bw() +
  labs(x = "", y = "Live weight\nmillion tonnes", color = "Source") +
  theme(
    legend.position = "bottom",
    legend.box.margin=margin(t = -25)
  )

### 2023-04-20 ms_results.Rmd
update file paths
Update results with new text

```{r load_data}
# Directory
datadir <- "outputs_20230330"

@@

Globalization describes the degree of international connectedness, which can be characterized by increasing flows of input, intermediate and final products among countries. Aquatic food exports increased `r round(100*((total_2019-total_1996)/total_1996),0)`%  (from `r total_1996` to `r total_2019` mil t; Fig 1a)  from 1996-2019. Over that period, both farmed and wild exports increased, though aquaculture grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%. Corresponding with the start of the COVID-19 pandemic, global aquatic food exports declined `r -1*round(100*((total_2020-total_2019)/total_2019), 0)`% overall in 2020 relative to 2019, with a `r -1*round(100*((total_capture_2020-total_capture_2019)/total_capture_2019),0)`% decline in capture exports, but a `r round(100*((total_aquaculture_2020-total_aquaculture_2019)/total_aquaculture_2019),1)`% increase in aquaculture exports (Fig 1a). Despite aquaculture comprising half of aquatic food production, capture fishery products constitute `r 100*round(total_capture_2019/total_2019, 1)`% of exports.

Another measure of the degree of globalization is the share of production exported. Domestic exports increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019, up from `r round(percent_export_total_dom_1996, 1)`% in 1996. while total exports reached 33% of all production in 2019. For comparison, the share of cereal production exported grew from around 10% in the late 1990s to 17% in the 2020s (Clapp 2022). Increasing marine capture export volumes despite stagnating production resulted in marine capture products having the greatest share of production destined for export (`r round(percent_marine_capture_exported_2019, 1)`% of production) and experienced the largest increases in the share exported, and largest increases in the share exported (Fig 2b). Aquaculture production more than doubled from 1996 to 2019, but aquaculture exports grew even faster, leading to an increasing share exported (Fig 2b). Despite increases, inland aquaculture still had the lowest share of production destined for export in 2019 (domestic exports represented only `r round(percent_inland_aqua_exported_2019, 1)`% of production) (Fig 1b). This finding clarifies prior debates about the orientation of aquaculture, while export trends suggest a need to consider international markets when crafting nutrition-sensitive policies (Golden et al. 2017; Belton et al 2018; Gephart et al. 2019).

Globalization exposes countries to external shocks, while also serving as a buffer against local shocks. Recent work exploring trade characteristics associated with systemic risk to shocks suggests higher exposure when networks are highly connected and concentrated, and when countries are more dependent on imports (Gephart et al. 2016;  Marchand et al. 2016; d’Amour et al. 2016). From 1996-2019, aquatic food trade became more connected with the average number of export partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`; Fig 4c). Marine capture networks are most highly connected, followed by marine aquaculture, with inland capture and aquaculture trade being the least connected.

Since 1996, aquatic food exports and imports have become moderately less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019. Compare this with crops where just 7 countries and the EU account for 90% of wheat exports and just four countries accounting >80% of maize exports (Clapp 2022). The declining concentration is driven by capture fishery exports, whereas both inland and marine aquaculture exports became somewhat more concentrated (Fig 2d). Aquaculture exports are generally concentrated among fewer countries than capture exports, corresponding to the high concentration of aquaculture production to a few key regions. Similarly, the concentration of trade for individual species tends to be much higher. The divergent trends in trade features and differences among aquatic food groups suggests differences in the degree and types of trade shock risks across the sector. Such differences were observed in responses to COVID-19 (Love et al. 2021). However, understanding risk to shocks across foods is a priority area for future work (Davis et al. 2021).

Though aquatic food production, distribution, and consumption remain highly uneven (Hicks et al. 2022), we found declinding import concentration, with `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 (Fig 2e). The more dispersed import patterns are likely associated with growing populations and expanding middle classes and urbanization, particularly in low- and middle-income countries, which are responsible for increasing aquatic food demand in many countries (Naylor et al. 2021; SOFIA 2022). However, the relationship between aquatic food demand and income varies across aquatic foods, with demand generally increasing with income for higher quality fish, but falling for lower quality fish (Naylor et al. 2021).

@@

Given the geographic patchiness of wild fish and aquaculture production, trade helps meet aquatic food demand in countries. Aquatic food imports are especially important where per capita demand is rising, aquaculture is limited, and wild fishery catch is stagnant. Corresponding to the geographical variability, we find the top importers, exporters, and bilateral flows to vary by habitat and farmed versus wild source, underscoring the importance of disaggregating trade (Fig S1-3). For example, Asia and Europe, and to a lesser extent, North America dominated marine capture and aquaculture trade networks in recent years (Fig S1). Since 1996, marine trade shifted toward Asia and Europe, with large increases in marine capture imports to and exports from Africa (Fig 2b).

At the country level, although some countries rank among the top traders across production methods, such as China for exports and the United Stated for imports, many countries are only top traders for one production method. Inland aquatic food trade is dominated by aquaculture, with the highest exports from Vietnam and China and highest imports by the United States, Japan, and South Korea (Fig S2-3). In general, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig S1, Fig S6).

Intraregional trade tends to be higher than interregional trade for all products due to shorter transport distances, historical ties, patterns of aquatic food preferences and established regional trade agreements (Natale et al. 2015). We find that this pattern generally holds for aquatic food trade as intraregional trade is the highest or second highest for all regions other than Oceania and South America. Since 1996, trade increased or remained approximately stable between nearly all regional trade pairs, other than within North America. At the country level, trade increased between `r round(percent_flow_increases, 1)`% of trade pairs. Despite trade increasing with partners across the globe, trade within Asia, Europe and Africa outpaced interregional trade with those regions. The largest average annual growth increases occurred for trade within Asia and within Europe, followed by trade between Europe and Asia (Fig 2a; Fig S6). Our trade estimates are ultimately from reported trade and therefore do not capture informal and unreported trade networks. Though estimated unreported trade is not globally available, it can be significant, especially for neighboring countries. For example, informal exports from Benin to Nigeria are estimated to be more than five times the formal exports (Benassi et al. 2019). Including informal trade would therefore likely strengthen intraregional trade patterns.

@@

habitatmethod_supply_trends_summary <- regional_supply_per_cap %>%
  select(year, habitat_method, supply_per_cap) %>%
  group_by(year, habitat_method) %>% 
  summarise(supply_per_cap = sum(supply_per_cap, na.rm = TRUE)) %>%
  filter(year %in% c(1996, 2019), habitat_method != "unknown") %>%
  pivot_wider(names_from = year, values_from = supply_per_cap, names_prefix = "year") %>%
  mutate(diff = year2019-year1996, 
         percent_diff = 100*diff/year1996)

@@

```{r}
national_supply_diversity <- read.csv(file.path(datadir, "fig4b_data.csv"))

n_import_exceeds_prod <- national_supply_diversity %>%
  filter(shannon_imports > shannon_prod) %>%
  nrow()

@@

Globally, annual aquatic food apparent consumption increased from `r signif(global_supply_1996, digit = 3)` kg per capita in 1996 to `r signif(global_supply_2019, digit = 3)` kg per capita in 2019. Our estimates are slightly lower than FAOSTAT, which reports global aquatic food consumption at 15.6 kg/capita/year in 1996 and 20.7 kg/capita/year in 2019. We found that aquatic food consumption increased across all regions outside of North America, which was relatively stable and South America, where aquatic food consumption declined `r -1*signif(regional_supply_trends_summary %>% filter(region == "South America") %>% pull(percent_diff), digit = 3)`% (Fig 3b). Global increase were driven by inland and marine aquaculture, which increased by `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019) and `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019), respectively. Meanwhile, inland capture consumption grew modestly (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland capture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland capture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019, a `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(percent_diff), digits = 3)`% increase), while marine capture consumption declined `r -1*signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019). Nevertheless, capture still makes up `r signif(cap_supply_percent, digits = 2)`% of global aquatic food consumption, with its contribution to regional aquatic food consumption ranging from `r signif(oceania_cap_supply_percent, digits = 2)`% in Oceania to `r signif(asia_cap_supply_percent, digits = 2)`% in Asia, where farmed consumption overtook wild consumption in `r asia_farmed_year`.

>>>>

```{r load_data}
# Directory
datadir <- "outputs"

@@

Globalization describes the degree of international connectedness, which can be characterized by increasing flows of input, intermediate and final products among countries. Aquatic food exports more than doubled from 1996-2019  (from `r total_1996` to `r total_2019` mil t; Fig 1a)  from 1996-2019. Over that period, both farmed and wild exports increased, though aquaculture grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%. Corresponding with the start of the COVID-19 pandemic, global aquatic food exports declined `r -1*round(100*((total_2020-total_2019)/total_2019), 0)`% overall in 2020 relative to 2019, with a `r -1*round(100*((total_capture_2020-total_capture_2019)/total_capture_2019),0)`% decline in capture exports, but a `r round(100*((total_aquaculture_2020-total_aquaculture_2019)/total_aquaculture_2019),1)`% increase in aquaculture exports (Fig 1a). Despite aquaculture comprising half of aquatic food production, capture fishery products constitute `r 100*round(total_capture_2019/total_2019, 1)`% of exports.

Another measure of the degree of globalization is the share of production exported. Domestic exports increased from `r round(percent_export_total_dom_1996, 1)`% to `r round(percent_export_total_dom_2019, 1)`% between 1996 and 2019, while total exports reached `r round(percent_export_total_2019, 1)`% of all production in 2019. For comparison, the share of cereal production exported grew from around 10% in the late 1990s to 17% in the 2020s (Clapp 2022). Increasing marine capture exports despite stagnating production resulted in marine capture products having the greatest share of production destined for export (`r round(percent_marine_capture_exported_2019, 1)`% of production) and the largest increases in the share exported (Fig 2b). Aquaculture production more than doubled from 1996 to 2019, but aquaculture exports grew even faster, increasing the share exported (Fig 2b). Despite increases, inland aquaculture still had the lowest share of production destined for export in 2019  (domestic exports represented only `r round(percent_inland_aqua_exported_2019, 1)`% of production) (Fig 1b). This finding clarifies standing debates about the orientation of aquaculture and export trends suggest a need to consider international markets when crafting nutrition-sensitive policies (Golden et al. 2017; Belton et al 2018; Gephart et al. 2019).

Globalization exposes countries to external shocks, while also serving as a buffer against local shocks. Recent work exploring trade characteristics associated with systemic risk to shocks suggests higher exposure when networks are highly connected and concentrated, and when countries are more dependent on imports (Gephart et al. 2016;  Marchand et al. 2016; d’Amour et al. 2016). By disaggregating aquatic food trade, we can evaluate the structural features of aquatic food trade associated with resilience to shocks. First, we found that aquatic food trade became more connected with the average number of export partners nearly doubling from 1996-2019  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`; Fig 1c). Marine capture networks are most highly connected, followed by marine aquaculture, with inland capture and aquaculture trade being the least connected.

Since 1996, aquatic food exports have become moderately less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019. Compare this with crops where just 7 countries and the EU account for 90% of wheat exports and just four countries accounting >80% of maize exports (Clapp 2022). Declining concentration is driven by capture fishery exports, whereas aquaculture exports became somewhat more concentrated (Fig 1d). Aquaculture exports are generally concentrated among fewer countries than capture exports, corresponding to the high concentration of aquaculture production to a few key regions. Similarly, the concentration of trade for individual species tends to be much higher. The divergent trends in trade features and differences among aquatic food groups suggests differences in the degree and types of trade shock risks across the sector. Such differences were observed in responses to COVID-19 (Love et al. 2021). However, understanding risk to shocks across foods is a priority area for future work (Davis et al. 2021).

Though aquatic food production, distribution, and consumption remain highly uneven (Hicks et al. 2022), we found declining import concentration, with `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 (Fig 2e). The more dispersed import patterns are likely associated with growing populations and expanding middle classes and urbanization, particularly in low- and middle-income countries, which are responsible for increasing aquatic food demand in many countries (Naylor et al. 2021; SOFIA 2022). However, the relationship between aquatic food demand and income varies across aquatic foods, with demand generally increasing with income for higher quality fish, but falling for lower quality fish (Naylor et al. 2021).

@@

Given the geographic patchiness of capture and aquaculture production, trade helps meet aquatic food demand in countries. Aquatic food imports are especially important where per capita demand is rising, aquaculture is limited, and wild fishery catch is stagnant. Corresponding to the geographical variability, we find the top importers, exporters, and bilateral flows to vary by habitat and farmed versus wild source, underscoring the importance of disaggregating trade (Fig S1-3). For example, Asia and Europe, and to a lesser extent, North America recently dominated marine capture and aquaculture trade networks whereas Asia dominates inland aquatic food trade  (Fig S1).

At the country level, although some countries rank among the top traders across all production methods, such as China for exports and the United Stated for imports, many countries are only top traders for one. China and Russia are the top marine capture aquatic food exporters, with China and the United States as the top importers (Fig S2-3). Meanwhile, Norway and Chile rank highest in marine aquaculture exports, with the United States and Japan leading imports (Fig S2-3). Inland aquatic food trade is dominated by aquaculture, with the highest exports from Vietnam and China and highest imports by the United States, Japan, and South Korea (Fig S2-3). In general, inland production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig S1, Fig S6).

Intraregional trade tends to be higher than interregional trade for all products due to shorter transport distances, historical ties, patterns of aquatic food preferences and established regional trade agreements (Natale et al. 2015). We find this pattern generally holds for aquatic food trade as intraregional trade is the highest or second highest for all regions other than Oceania and South America (Fig 2b). Since 1996, trade increased or remained approximately stable between nearly all regional trade pairs, other than within North America (Fig S6). At the country level, trade increased between two thirds of trade pairs. Despite trade increasing with partners across the globe, trade within Asia, Europe and Africa grew faster. The largest average annual growth increases occurred for trade within Asia and Europe, followed by trade between Europe and Asia (Fig 2b; Fig S6). Our trade estimates are ultimately from reported trade and therefore do not capture informal and unreported trade networks. Though estimated unreported trade is not globally available, it can be significant, especially for neighboring countries. For example, informal exports from Benin to Nigeria are estimated to be more than five times the formal exports (Benassi et al. 2019). Including informal trade would therefore likely strengthen intraregional trade patterns.

@@

habitatmethod_supply_trends_summary <- regional_supply_per_cap %>%
  select(year, habitat_method, supply, pop) %>%
  group_by(year, habitat_method) %>% 
  summarise(supply = sum(supply, na.rm = TRUE),
            pop = sum(pop, na.rm = TRUE)) %>%
  mutate(supply_per_cap = 1000 * supply/pop) %>%
  select(-supply, -pop) %>%
  filter(year %in% c(1996, 2019), habitat_method != "unknown") %>%
  pivot_wider(names_from = year, values_from = supply_per_cap, names_prefix = "year") %>%
  mutate(diff = year2019-year1996, 
         percent_diff = 100*diff/year1996)

@@

@@

Globally, annual aquatic food apparent consumption increased from `r signif(global_supply_1996, digit = 3)` kg per capita in 1996 to `r signif(global_supply_2019, digit = 3)` kg per capita in 2019 (Fig 3a). Our estimates are slightly lower than FAOSTAT, which reports global aquatic food consumption at 15.6 kg/capita/year in 1996 and 20.7 kg/capita/year in 2019. We found aquatic food consumption increased across all regions outside of North America, which was relatively stable, and South America, where aquatic food consumption declined `r -1*signif(regional_supply_trends_summary %>% filter(region == "South America") %>% pull(percent_diff), digit = 3)`% (Fig 3b). Global increases were driven by inland and marine aquaculture, which increased by `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019) and `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019), respectively. Meanwhile, inland capture consumption grew from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland capture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland capture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019, while marine capture consumption declined `r -1*signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019). Nevertheless, capture still makes up `r signif(cap_supply_percent, digits = 2)`% of global aquatic food consumption, with its contribution to regional aquatic food consumption ranging from `r signif(oceania_cap_supply_percent, digits = 2)`% in Oceania to `r signif(asia_cap_supply_percent, digits = 2)`% in Asia, where farmed consumption overtook wild consumption in `r asia_farmed_year`.

### 2023-04-20 supply_diversity.Rmd
Update results with new text

>>>>

# Methods

## Estimating supply diversity
	We separately calculated the Shannon diversity for the overall supply, the diversity of supply sourced from domestic products, and the supply sourced from foreign products. Additionally, we calculated the Shannon diversity of domestic production, imports, and exports. One notable limitation of applying this measure is that instances where production or trade is only estimated at the genus or higher level will be interpreted in the calculation as a single species. Consequently, countries with low species resolution will generally appear to have less diverse production, limiting the ability to compare supply diversity across countries. To address this issue, we first matched scientific groups to species based on data in FishBase50 and SealifeBase53. We then calculated the global production distribution within each phylogenetic group based on a lognormal distribution of produced species. Proportions for each potential species within a phylogenetic group were derived from the lognormal global distribution. These proportions were matched to species in two ways. For species that had been previously reported by a country they were matched in order by the volume reported for that year and reported species were followed by the unreported species. This meant that, for a given year, the highest proportion derived from the lognormal distribution was matched to the species that the country had reported the largest volume for
